Confluence in Lens Synthesis

### 2020-04-17 papers/BijectiveLenses/IWC/main.tex
update abstract

\begin{abstract}
  Bidirectional programs are programs where a single expression has multiple
  interpretations, encoding a number of functions with a single piece of syntax.
  Lenses are bidirectional programs with semantic guarantees, and lens languages
  have been developed for synchronizing string formats, building updatable views
  in databases, and more. These languages often are typed, where a well-typed
  expression will provide the semantic guarantees. Recent research has shown how
  string lenses can be synthesized from their types, pairs of regular
  expressions. However, providing completeness guarantees is quite tricky on
  these languages, due in large part to the many equivalences on regular
  expressions.

  In proving a completeness guarantee, a confluence-like property,
  $R$-confluence, must be shown. This property can be thought of as a
  generalization of standard confluence. Where standard confluence is
  parameterized by a set $S$ and a binary relation $\rightarrow$, indexed
  confluence is parameterized by an additional binary relation $R \subseteq S
  \times S$. In this model, confluence is the specific case where $R$ is just
  equality. We show how the proof that the transitive closure of a rewrite
  system is $R$-confluent differs from the proof strategy for standard
  confluence.
\end{abstract}

@@

To make bidirectional programming easier, Optician~\cite{optician} synthesizes
bidirectional programs from these types, specifically in Boomerang, a language
for synchronizing string data. Optician takes as input two regular expressions
(which serve as the type of a Boomerang lens) and a set of examples specifying
input-output behavior, and synthesizes a well-typed lens between the languages
of those regular expressions.

The judgement $\Lens : \Regex \Leftrightarrow \RegexAlt$ means that $\Lens$
bidirectionally converts between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$. For synthesis, given $\Regex$ and $\RegexAlt$, we must
find a lens $\Lens$ such that $\Lens : \Regex \Leftrightarrow \RegexAlt$. There
are a number of typing rules for lenses, but the rules for type-equivalence and
composition (shown below) do not permit an efficient search algorithm. Note that
in type equivalence we use a finer notion of equivalence than language
equivalence, we instead use star-semiring equivalence, denoted $\SSREquiv$. We were unable to make
our proof strategy work for language equivalence, so instead guarantee
completeness only on the subset of lenses well-typed with star-semiring
equivalence.
\begin{mathpar}
  \inferrule[\RewriteRegexLensRule{}]
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \SSREquiv \Regex'\\
    \RegexAlt \SSREquiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }

>>>>

\begin{abstract}
  A lens is a program that can be executed both forwards and backwards, from
  input to output and from output back to input again. Domain-specific languages
  for defining lenses have been developed to help users synchronize text files,
  and construct different ``views'' of databases, among other applications.
  Recent research has shown how string lenses can be synthesized from their
  types, which are pairs of regular expressions. However, guaranteeing that we
  can synthesize all possible lenses is quite tricky on these languages, due in
  large part to the many equivalences on regular expressions.

  Our proof that we can synthesize all string lenses involves proving a
  confluence-like property, indexed by a an additional binary relation $R$. In
  this model, standard confluence is the specific case where $R$ is equality. We
  show how our proof that the transitive closure of a rewrite system is
  $R$-confluent differs from the proof strategy for standard confluence.
\end{abstract}

@@

Optician~\cite{optician}, an extension of Boomerang, makes bidirectional
programming easier by supporting synthesis of bidirectional string
transformations. More specifically, it takes as input two regular expressions
($\Regex$ and $\RegexAlt$, which serve as the type of a Boomerang lens) and a set of examples specifying
input-output behavior, and synthesizes a well-typed lens between the languages
of those regular expressions.

To indicate a lens $\Lens$ is well typed, and converts between the languages of
$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$ For synthesis, given $\Regex$ and $\RegexAlt$, we must
find a lens $\Lens$ such that $\Lens : \Regex \Leftrightarrow \RegexAlt$. There
are a number of typing rules for lenses, but the rules for type-equivalence and
composition (shown below) do not permit an efficient search algorithm. Note that
in type equivalence we use a finer notion of equivalence than language
equivalence, we instead use star-semiring equivalence, denoted $\SSREquiv$. We were unable to make
our proof strategy work for language equivalence, so instead guarantee
completeness only on the subset of lenses well-typed with star-semiring
equivalence.
\begin{mathpar}
  \inferrule[\RewriteRegexLensRule{}]
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \SSREquiv \Regex'\\
    \RegexAlt \SSREquiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }
